\TitlePt

### 2016-03-19 abstract.tex
Portuguese version of the abstract

Apresenta-se, nesta tese, ...

\end{abstract}

@@

We present methodological improvements to forward modeling and large-scale
inversion of satellite gravity data.
For this purpose, we developed two open-source software projects.
The first, is a C language suite of command-line programs called
\textit{Tesseroids}.
The programs calculate the gravitational potential, acceleration, and gradient
tensor of a spherical prism, or tesseroid.
\textit{Tesseroids} implements and extends an adaptive discretization algorithm
to automatically ensure the accuracy of the computations.
Our numerical experiments show that, to achieve the same level of accuracy, the
gravitational acceleration components require finner discretization than the
potential.
In turn, the gradient tensor requires finner discretization still than the
acceleration.
The second open-source project is \textit{Fatiando a Terra}, a Python language
library for inversion, forward modeling, data processing, and visualization.
The library allows the user to combine the forward modeling and inversion tools
to implement new inversion methods.
The gravity forward modeling tools include an implementation of the
\textit{Tesseroids} algorithm.
We combined the inversion and tesseroid forward modeling utilities of
\textit{Fatiando a Terra} to develop a new method for fast non-linear gravity
inversion.
The method estimates the depth of the crust-mantle interface (the Moho) based
on observed gravity data using a spherical Earth approximation.
We extended the computationally efficient Bott's method to include smoothness
regularization and use tesseroids instead right rectangular prisms.
The inversion is controlled by three hyper-parameters: the regularization
parameter, the Moho density-contrast, and the depth of the reference Moho.
We employ two cross-validation procedures to automatically estimate these
parameters.
Tests on synthetic data confirm the capability of the proposed method to
estimate smoothly varying Moho depths and the three hyper-parameters.
Finally, we applied our inversion method to produce a Moho depth model for
South America.
Our model fits the gravity data and seismological Moho depth estimates
in the oceanic areas and the central and eastern portions of the continent.
We observe large misfits in the Andes region, where Moho depth is largest.
In Amazon, Solimões, and Paraná Basins, the model fits the observed gravity
but disagrees with seismological estimates.
These discrepancies suggest the existence of density-anomalies in the crust or
upper mantle, as has been suggested in the literature.
\end{foreignabstract}

>>>>

Nós apresentamos avanços metodológicos na área de modelagem direta e inversão
regional de dados de gravimetria por satélite.
Com esse fim, nós desenvolvemos dois projetos de software de código livre.
O primeiro é um conjunto de programas de linha de comando feitos na linguagem C
chamado \textit{Tesseroids}.
Os programas calculam o potencial, aceleração e tensor gradiente gravitacional
de um prisma esférico, ou tesseroide.
\textit{Tesseroids} implementa e aprimora um algoritmo de discretização
adaptativa para garantir a acurácia das computações automaticamente.
Nossos resultados com testes numéricos mostram que, para obter o mesmo nível de
acurácia, a aceleração gravitacional demanda uma discretização mais fina que o
potencial.
Por sua vez, o tensor gradiente gravitacional demanda discretização mais final
ainda que para a aceleração.
O segundo software é o \textit{Fatiando a Terra}, uma biblioteca feita na
linguagem Python para inversão, modelagem direta, processamento e visualização
de dados.
A biblioteca permite que o usuário combine as ferramentas de modelagem direta e
inversão para implementar novos métodos de inversão.
As ferramentas de modelagem direta incluem uma implementação do algoritmo
utilizado no programa \textit{Tesseroids}.
Nós combinamos os recursos de inversão e modelagem direta com tesseroides do
\textit{Fatiando a Terra} para desenvolver um método rápido para a inversão
não-linear de dados de gravidade.
O método estima a profundidade da interface crosta-manto (a Moho) baseado em
dados de gravidade utilizando uma aproximação esférica para a Terra.
Nós adaptamos o método de Bott, que é computacionalmente eficiente, para
incluir regularização de suavidade  e utilizar tesseroides ao invés de prismas
retangulares retos.
A inversão é controlada por três hiper-parâmetros: o parâmetro de
regularização, o contraste de densidade na Moho e a profundidade da Moho de
referência.
Nós aplicamos dois tipos de validação cruzada para estimar esses parâmetros de
maneira automática.
Testes com dados sintéticos confirmam a capacidade do método proposto de
estimar os três hiper-parâmetros e o relevo suave da Moho.
Por fim, nós aplicamos nosso método de inversão para gerar um modelo de
profundidades da Moho para a América do Sul.
Nosso modelo ajusta os dados de gravidade observados e estimativas da
profundidade da Moho provenientes da sismologia
nas regiões oceânicas e nas partes central e Leste do continente.
Nós observamos desajustes aos dados na região dos Andes, onde a profundidade da
Moho é a maior para o continente.
Nas bacias do Amazonas, Solimões e Paraná, o modelo ajusta os dados de
gravidade mas não as estimativas da sismologia.
Essas discrepâncias indicam a presença de anomalias de densidade na crosta ou
manto superior, como sugerido anteriormente na literatura.
\end{abstract}

@@

We present methodological improvements to forward modeling and regional
inversion of satellite gravity data.
For this purpose, we developed two open-source software projects.
The first is a C language suite of command-line programs called
\textit{Tesseroids}.
The programs calculate the gravitational potential, acceleration, and gradient
tensor of a spherical prism, or tesseroid.
\textit{Tesseroids} implements and extends an adaptive discretization algorithm
to automatically ensure the accuracy of the computations.
Our numerical experiments show that, to achieve the same level of accuracy, the
gravitational acceleration components require finner discretization than the
potential.
In turn, the gradient tensor requires finner discretization still than the
acceleration.
The second open-source project is \textit{Fatiando a Terra}, a Python language
library for inversion, forward modeling, data processing, and visualization.
The library allows the user to combine the forward modeling and inversion tools
to implement new inversion methods.
The gravity forward modeling tools include an implementation of the
algorithm used in the \textit{Tesseroids} software.
We combined the inversion and tesseroid forward modeling utilities of
\textit{Fatiando a Terra} to develop a new method for fast non-linear gravity
inversion.
The method estimates the depth of the crust-mantle interface (the Moho) based
on observed gravity data using a spherical Earth approximation.
We extended the computationally efficient Bott's method to include smoothness
regularization and use tesseroids instead right rectangular prisms.
The inversion is controlled by three hyper-parameters: the regularization
parameter, the Moho density-contrast, and the depth of the reference Moho.
We employ two cross-validation procedures to automatically estimate these
parameters.
Tests on synthetic data confirm the capability of the proposed method to
estimate smoothly varying Moho depths and the three hyper-parameters.
Finally, we applied our inversion method to produce a Moho depth model for
South America.
Our model fits the gravity data and seismological Moho depth estimates
in the oceanic areas and the central and eastern portions of the continent.
We observe large misfits in the Andes region, where Moho depth is largest.
In Amazon, Solimões, and Paraná Basins, the model fits the observed gravity
but disagrees with seismological estimates.
These discrepancies suggest the existence of density-anomalies in the crust or
upper mantle, as has been suggested in the literature.
\end{foreignabstract}
